---
title: "hw2"
author: "jing wen-jw4059"
date: "2/8/2021"
output: pdf_document
---



```{r}
# Define output.image function (Lab 2)
output.image<-function(vector) {
  digit<-matrix(vector, nrow=16, ncol=16)
  index= seq(from=16, to =1, by=-1)
  sym_digit = digit[,index]
  image(sym_digit, col= gray((8:0)/8), axes=FALSE)
}

# Read in digits features for 3,5,8
digit_3 <- read.table("train_3.txt",header = FALSE,sep=',')
digit_5 <- read.table("train_5.txt",header = FALSE,sep=',')
digit_8 <- read.table("train_8.txt",header = FALSE,sep=',')

# Define temporary data matrix
X_temp <- rbind(digit_3,digit_5,digit_8)
dim(X_temp)

# Extract "test" cases
# Test case 1
ConstructCase_1 <- X_temp[20,] 
output.image(as.matrix(ConstructCase_1))
ConstructCase_1 <- unlist(ConstructCase_1) # Not needed but might be helpful

# Test case 2
ConstructCase_2 <- X_temp[735,] 
output.image(as.matrix(ConstructCase_2))
ConstructCase_2 <- unlist(ConstructCase_2) # Not needed but might be helpful

# Test case 3
ConstructCase_3 <- X_temp[1260,] 
output.image(as.matrix(ConstructCase_3))
ConstructCase_3 <- unlist(ConstructCase_3) # Not needed but might be helpful

# Remove cases 20, 735, 1260 from original dataframe
X <- X_temp[-c(20,735,1260),]
dim(X)
```
Q4
```{r}
#1
zip=read.table("zip_test.txt",header =FALSE,sep='')
zip.test=zip[zip$V1=="3"|zip$V1=="5"|zip$V1=="8",]
digit_3$response="3"
digit_5$response="5"
digit_8$response="8"
X_temp.2 <- rbind(digit_3,digit_5,digit_8)
response=X_temp.2$response
list2=colnames(zip.test[,1:256])
zip.t=zip.test[,2:ncol(zip.test)]
names(zip.t)=list2
library(MASS)
model <- lda(X_temp,grouping=response)
model.pre1=predict(model, zip.t)
lda.class1=model.pre1$class
mean(lda.class1 != zip.test$V1) #test error
model.pre2=predict(model, X_temp)
lda.class2=model.pre2$class
mean(lda.class2 != response) #train error
#2
pca49<-prcomp(X_temp, rank. = 49)
response=as.numeric(response)
x.49<-cbind(response, pca49$x)
x.49<-data.frame(x.49)
t.49<-cbind(response=zip.test[,1], predict(pca49, zip.t))
t.49<-data.frame(t.49)
model2 <- lda(x.49[,-1],grouping=response)
model.pre3=predict(model2, t.49[,-1])
lda.class3=model.pre3$class
mean(lda.class3 != t.49$response) #test error
model.pre4=predict(model2, x.49[,-1])
lda.class4=model.pre4$class
mean(lda.class4 != x.49$response) #train error
#3
library(nnet)
mlr=multinom(response~.,data=x.49)
model.pre5=predict(mlr, t.49[,-1])
mean(model.pre5 !=  t.49[,1]) #test error
model.pre6=predict(mlr, x.49[,-1])
mean(model.pre6 != x.49[,1]) #train error
```

q5
```{r}
library(plyr)
library(quantmod)
stock=function(x){
 data=getSymbols(x, auto.assign = FALSE, from ="2020-01-01", to = "2021-01-01") 
 return(data)
}
list=c("AXP","AMGN","AAPL","BA","CAT","CVX","CSCO","KO","DIS","DOW","GS","HD","HON","INTC","JNJ","JPM","MCD","MRK","MSFT","NKE","PG","CRM","TRV","VZ","V","WBA","WMT","IBM","MMM","UNH")
dataf=llply(list,stock)
dataf=as.data.frame(matrix(unlist(dataf), nrow=length(unlist(dataf[1]))))
colnames(dataf)=list
#2
pca=prcomp(dataf)
biplot(pca)
#3
dataf.C <- scale(dataf)
pca.C <- prcomp(dataf.C)
biplot(pca.C)
#4
ret=c()
ret.r=function(x){
  ret=diff(x)/x[1:(length(x)-1)]
  return(ret)
}
dataf.r=lapply(dataf,ret.r)
dataf.r=as.data.frame(matrix(unlist(dataf.r), nrow=length(unlist(dataf.r[1]))))
colnames(dataf.r)=list
dataf.C.r <- scale(dataf.r)
pca.C.r <- prcomp(t(dataf.C.r))
screeplot(pca.C.r,type="lines",col="violet")
CPVE <- (cumsum((pca.C.r$sdev)^2)/sum((pca.C.r$sdev)^2))
plot(1:30,CPVE,type="l",col="blue",
     ylim=c(.5,1),xlab="Principal Component",ylab="Cumulative PVE")
points(1:30,CPVE,col="blue")
```
#2 from the biplot, stock AAF seems to have a larger eigen value than the other stocks.AAF is the only that fully show up other than the other stocks in the plot.
#3, after scale the variables, CRM turns out to be important from the biplot
#4 from screeplot, it seems PC1 caputres the most variation in PCA.for returns, all the stocks have the smiliar direction.If each stock were fluctuating up and down randomly and independent of all the other stocks,i would expect scree plot to be a sraight line.
q6
```{r}
x.c <- scale(X,center=TRUE,scale=FALSE)
x.pca=prcomp(x.c)
biplot(x.pca)
screeplot(x.pca,type="l")
#2
PVE <- (cumsum((x.pca$sdev)^2)/sum((x.pca$sdev)^2))
x=c(1:256)
plot(x,PVE)
sum(PVE<0.9)
#3
xr=x.pca$rotation
par(mfrow=c(4,4),pin = c(1,1))
for (i in 1:16){
  output.image(as.matrix(xr[,i]))
}

#4
list=c(3,58,256)
construct1 <- rep(0,3)
construct2 <- rep(0,58)
construct3 <- rep(0,256)
for (j in list){
   for (i in 1:j) {
  construct1 <- construct1 + sum(x.pca$rotation[,i]*ConstructCase_1)*x.pca$rotation[,i]
   }
   output.image(construct1)
}

for (j in list){
   for (i in 1:j) {
  construct2 <- construct2 + sum(x.pca$rotation[,i]*ConstructCase_2)*x.pca$rotation[,i]
   }
   output.image(construct2)
}

for (j in list){
   for (i in 1:j) {
  construct3 <- construct3 + sum(x.pca$rotation[,i]*ConstructCase_3)*x.pca$rotation[,i]
   }
   output.image(construct3)
}
 

```